Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-11-13 paper/main.tex
minor

This paper presents a new compositional framework for automatically
deriving resource bounds on C programs.  This new
approach is an attempt to unify the two aforementioned
categories: It solves the compositionality issues of techniques
for numerical imperative code by adapting amortized-analysis--based
techniques from the functional world.
%
Our automated analysis is able to infer
resource bounds on C programs with mutually-recursive functions and integer loops.
%
The resource behavior of functions can be summarized
in a function specification that can be used at every
call site without accessing the function body.
%
To our knowledge this is the first technique based on
amortized analysis that is able to derive bounds that depend on negative numbers
and differences of variables.  It is also the first
resource analysis technique for C that deals naturally with
recursive functions and sequenced loops, and can handle resources that can
become available during execution (e.g., when freeing
memory).  Compared to more classical
approaches based on ranking functions, our tool inherits
the benefits of amortized reasoning.  Using only one
simple mechanism, it handles:
\itemskip
\begin{itemize}
\item interactions between \emph{sequential loops or
  function calls} through size changes of variables,
\itemskipIn
\item \emph{nested loops} that influence each other
  with the same set of modified variables,
\itemskipIn
\item and \emph{amortized bounds} as found, for example, in
  the Knuth-Morris-Pratt algorithm for string search.
\end{itemize}
\itemskip
The main innovation that makes amortized analysis work
on imperative languages is to base the analysis on a
Hoare-like logic and to track multivariate quantities instead
of program variables.  This leads to precise bounds
expressed as functions of sizes $|[x, y]| = \max(0, y-x)$ of
intervals.
%
A distinctive feature of our analysis system is that
it reduces linear bound inference  to a linear optimization problem
that can solved by off-the-shelf LP solvers.  This enables the
efficient inference of global bounds for larger programs.
%
Moreover, our local inference rules automatically generate
proof certificates that can be easily checked in linear time.

>>>>

This paper presents a new compositional framework for automatically
deriving resource bounds on C programs.  This new
approach is an attempt to unify the two aforementioned
categories: It solves the compositionality issues of techniques
for numerical imperative code by adapting amortized-analysis--based
techniques from the functional world.
%
Our automated analysis is able to infer
resource bounds on C programs with mutually-recursive functions and integer loops.
%
The resource behavior of functions can be summarized
in a function specification that can be used at every
call site without accessing the function body.
%
To our knowledge this is the first technique based on
amortized analysis that is able to derive bounds that depend on negative numbers
and differences of variables.  It is also the first
resource analysis technique for C that deals naturally with
recursive functions and sequenced loops, and can handle resources that can
become available during execution (e.g., when freeing
memory).  Compared to more classical
approaches based on ranking functions, our tool inherits
the benefits of amortized reasoning.  Using only one
simple mechanism, it handles:
\itemskip
\begin{itemize}
\item interactions between \emph{sequential loops or
  function calls} through size changes of variables,
\itemskipIn
\item \emph{nested loops} that influence each other
  with the same set of modified variables,
\itemskipIn
\item and \emph{amortized bounds} as found, for example, in
  the Knuth-Morris-Pratt algorithm for string search.
\end{itemize}
\itemskip
The main innovation that makes amortized analysis work
on imperative languages is to base the analysis on a
Hoare-like logic and to track multivariate quantities instead
of program variables.  This leads to precise bounds
expressed as functions of sizes $|[x, y]| = \max(0, y-x)$ of
intervals.
%
A distinctive feature of our analysis system is that
it reduces linear bound inference  to a linear optimization problem
that can be solved by off-the-shelf LP solvers.  This enables the
efficient inference of global bounds for larger programs.
%
Moreover, our local inference rules automatically generate
proof certificates that can be easily checked in linear time.
